### 2016-05-25 Projects/DelCo_2014_2015/Reports/Propane_Gas_Fire_Experiments/TeX_Files/Propane_Gas_Fire_Experiments.tex
Minor edits to gas burner report

\section{Instrumentation}
\label{sec:Instrumentation}
The structures were instrumented for temperature, gas velocity, heat flux, and gas concentration measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single thermocouples were installed in conjunction with bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple wire was protected with an 3.2~mm (0.13~in) diameter inconel sheath. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. Calibrated pumps pulled gas samples through a sample conditioning system to eliminate moisture in the sample. Then, the dry gas sample was piped to a series of gas analyzers and the concentrations of oxygen, carbon monoxide, and carbon dioxide were measured. A legend is presented in Fig.~\ref{fig:Instrumentation_Legend} to clarify the instrumentation schematics presented in the follow sections.

@@

Three propane burners, pictured in Fig.~\ref{fig:burners}, were used as the fuel source in each experiment. Each burner had a square opening of side length 0.6~m (2~ft) located 0.14~m (5.5~in) above the floor and were positioned 0.6~m (2~ft) from the south and west walls on the first floor of each structure. Propane was supplied to the burners at a measured flow rate during all experiments.
% INCLUDE HRR CALCULATIONS HERE

\begin{figure}[!ht]
	\includegraphics[width=0.9\columnwidth]{../Figures/Pictures/burners}
	\caption[Three propane burners.]{Three propane burners located 0.6~m (2~ft) off the south and west interior walls of the first floor provided the fire source for the experiments.}
	\label{fig:burners}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Structure_Dimensioned_Instrumentation}
	\caption{Location of instrumentation in the East Structure.}
	\label{fig:east_instrumentation}
\end{figure}
\FloatBarrier

Each bare-bead thermocouple array was composed of eight thermocouples. Three bi-directional probe and solid thermocouple arrays were centered in each exterior doorway of the structure and contained eight probes as shown in Fig.~\ref{fig:BDP_arrays}. The fourth bi-directional probe and solid thermocouple array, also presented in Fig.~\ref{fig:BDP_arrays}, was located at the opening of the roof vent 0.32~m (12.75 in) above the room ceiling and contained three probes centered between the east and west sides of the vent. The position of each probe and thermocouple pair relative to the south wall of the vent are listed in Table~\ref{table:east_channel_list}. The total heat flux gauge/radiometer pairs were set to be aimed to view the ceiling. Gas concentrations were pulled through 9.5~mm (0.375~in) diameter stainless steel tubing. The height of each individual sensor in the sensor arrays are listed in the channel list found in Table~\ref{table:east_channel_list}.

\begin{figure}[!ht]
	\includegraphics[width=0.45\columnwidth]{../Figures/Pictures/doorway_BDPs}
	\\~\\
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/roof_vent_BDPs.png}
	\caption[Bi-directional probe plus solid thermocouple arrays in East Structure.]{Bi-directional probe plus solid thermocouple array in south exterior doorway (top) and in roof vent (bottom) of the East Structure.}
	\label{fig:BDP_arrays}
\end{figure}

@@

\subsection{West Structure}
The first floor of the West Structure was instrumented with three bare-bead thermocouple arrays, two bi-directional probe plus solid thermocouple arrays, and one gas sample inlet pipe. The second floor was equipped with three bare-bead thermocouple arrays, four bi-directional probe plus solid thermocouple arrays, two total heat flux sensor pairs (one facing horizontal and one facing vertical), and one gas sample inlet pipe. The location of the instrumentation in the West Structure is shown in Fig.~\ref{fig:west_instrumentation}.

\begin{figure}[!ht]
	\includegraphics[width=0.94\columnwidth]{../Figures/Floor_Plans/West_Structure_2nd_Floor_Dimensioned_Instrumentation}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_1st_Floor_Dimensioned_Instrumentation}
	\caption[Location of instrumentation in the West Structure.]{Location of instrumentation in the second floor (top) and first floor (bottom) of the West Structure.}
	\label{fig:west_instrumentation}
\end{figure}

Every thermocouple array contained eight bare-bead thermocouples, each bi-directional probe and solid thermocouple array contained 8 probes, and gas concentrations were pulled through 9.5~mm (0.375~in) diameter stainless steel tubing located 1.2~m (4~ft) above the floor.
\clearpage

@@

\subsection{Measurement Uncertainty}
\label{sec:Uncertainty}
There are different components of uncertainty in the reported length, mass, temperature, heat flux, gas concentration, differential pressure, and gas velocity measurements. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

\subsubsection{Bi-Directional Probes}
Bi-directional probes and single thermocouples were used to measure the velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. A single bare-bead Chromel-Alumel (type K) thermocouple with a 1.0~mm (0.04~in) nominal diameter was co-located with each probe. The thermocouple wire was protected with a 3.2~mm (0.125~in) diameter inconel sheath. A gas velocity measurement study examining the doorway flow of pre-flashover compartment fires yielded expanded uncertainty measurements ranging from $\pm$0.14 to $\pm$0.22 for bi-directional probes similar to the ones described here~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments was estimated to be $\pm$18\%.

\subsubsection{Gas Flow Meter}
A rotary type gas meter was used to measure the flow of propane to the gas burners. The manufacturer of the meter quantifies the maximum error as 2\%.

@@

% ==========================
% = EXPERIMENTAL PROCEDURE =
% ==========================
\chapter{Experimental Procedure}
\label{chap:Experimental_Procedure}
A similar procedure was followed for all experiments described in this report. First, three propane burners were ignited in sequential order. Next, different doors and vents in the structure were opened and closed to change the ventilation pattern within the structure. Finally, the burners were turned off in sequential order and the fire was extinguished.

\section{East Structure Tests}
\label{sec:east_procedure}
Five different tests, Tests 2--6, were conducted in the East Structure. Tests~2--4 followed an identical order of events. Fig.~\ref{fig:Tests_2-4_layout} includes a floor plan schematic and table of event times corresponding to the data files for Tests~2--4. The procedures for Test~5 and Test~6 are outlined in a similar manner in Fig.~\ref{fig:east_test_5} and Fig.~\ref{fig:east_test_6}, respectively. Tests~5 and 6 involved performing a specific set of events three times sequentially. To avoid repetition of identical actions in the event time table, each set of events has been denoted as a ``sequence'' (abbreviated as ``seq.'' in the table). Additionally, Table~\ref{table:HRR_East} lists the calculated heat release rate for each test and sequence.

\begin{table}[!ht]
\caption{Calculated Heat Release Rate for East Structure Tests}
\begin{tabular}{lccccc}
 \toprule
 					& 	\multicolumn{5}{c}{\textbf{Heat Release Rate (kW)}} \\
\textbf{Test} 		&	\begin{tabular}{c} Corner \\ Burner On \end{tabular} 
					&	\begin{tabular}{c} Middle \\ Burner On \end{tabular}
					&	\begin{tabular}{c} Center \\ Burner On \end{tabular}
					& 	\begin{tabular}{c} Center \\ Burner Off \end{tabular}
					&   \begin{tabular}{c} Middle \\ Burner Off \end{tabular} \\
 \midrule
Test 2$^*$ 			& 338 & 651	& 901 & 634 & 330 \\
Test 3$^*$			& 326 & 650	& 905 & 631 & 326 \\
Test 4				& 351 & 654	& 952 & 652 & 337 \\
Test 5 - Seq. 1		& N/A & N/A	& 725 & N/A & N/A \\
Test 5 - Seq. 2		& N/A & N/A	& 728 & N/A & N/A \\
Test 5 - Seq. 3		& N/A & N/A	& 723 & N/A & N/A \\
Test 6 - Seq. 1		& N/A & N/A	& 722 & N/A & N/A \\
Test 6 - Seq. 2		& N/A & N/A	& 721 & N/A & N/A \\
Test 6 - Seq. 3		& N/A & N/A	& 714 & N/A & N/A \\
\bottomrule
\end{tabular}
\label{table:HRR_East}
\end{table}

@@

\section{West Structure Tests}
\label{sec:west_procedure}
Four different tests---Tests~22, 23, 24, and 25---were conducted in the West Structure. Tests~22 and 23 followed an identical procedure to change the ventilation. Fig.~\ref{fig:west_test_22} includes a floor plan schematic and table of event times corresponding to the data files for Tests~22 and 23. Similarly, Fig.~\ref{fig:west_test_24} includes a floor plan schematic and table of event times that outline the procedure followed during Tests~24 and 25.

>>>>

\section{Instrumentation}
\label{sec:Instrumentation}
The structures were instrumented for temperature, gas velocity, heat flux, and gas concentration measurements. Gas temperatures in the burn rooms were measured with bare-bead, Chromel-Alumel (type K) thermocouples. Additional single thermocouples were installed in conjunction with bi-directional probes for gas velocity measurements. The single thermocouples were bare-bead, Chromel-Alumel (type K) thermocouples with a 1.0~mm (0.04~in) nominal diameter. The thermocouple wire was protected with a 3.2~mm (0.13~in) diameter inconel sheath. Schmidt-Boelter gauges were used to measure both total heat flux and radiant heat flux (radiometer). A radiometer is a total heat flux gauge with a zirconium plate to prevent contributions from convective heat transfer. Calibrated pumps pulled gas samples through a sample conditioning system to eliminate moisture in the sample. Then, the dry gas sample was piped to a series of gas analyzers and the concentrations of oxygen, carbon monoxide, and carbon dioxide were measured. A legend is presented in Fig.~\ref{fig:Instrumentation_Legend} to clarify the instrumentation schematics presented in the follow sections.

@@

Three propane burners, pictured in Fig.~\ref{fig:burners}, were used as the fuel source in each experiment. Each burner had a square opening of side length 0.6~m (2~ft) located 0.14~m (5.5~in) above the floor and were positioned 0.6~m (2~ft) from the south and west walls on the first floor of each structure. Propane was supplied to the burners during all experiments. The flow of propane to each burner was controlled by a high-precision turn valve, and the total displaced gas volume was measured using a rotary gas meter.
% INCLUDE HRR CALCULATIONS HERE

\begin{figure}[!ht]
	\includegraphics[width=0.9\columnwidth]{../Figures/Pictures/burners}
	\caption[Three propane burners used as fire source.]{Three propane burners located 0.6~m (2~ft) off the south and west interior walls on the first floor provided the fire source for the experiments.}
	\label{fig:burners}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/East_Structure_Dimensioned_Instrumentation}
	\caption{Locations and labels of instrumentation in the East Structure.}
	\label{fig:east_instrumentation}
\end{figure}
\FloatBarrier

Each bare-bead thermocouple array was composed of eight thermocouples. Three bi-directional probe and solid thermocouple arrays (A7, A8, and A9) were centered in each exterior doorway of the structure and contained eight probes as shown in Fig.~\ref{fig:BDP_arrays}. The fourth bi-directional probe and solid thermocouple array (A10), also presented in Fig.~\ref{fig:BDP_arrays}, was located at the opening of the roof vent and contained three probes centered between the east and west sides of the vent. The position of each probe and thermocouple pair relative to the south wall of the vent are listed in Table~\ref{table:east_channel_list}. The total heat flux gauge/radiometer pairs were set to be aimed to view the ceiling. Gas concentrations were pulled through 9.5~mm (0.38~in) diameter stainless steel tubing. The height of each individual sensor in the sensor arrays are listed in the channel list found in Table~\ref{table:east_channel_list}.

\begin{figure}[!ht]
	\includegraphics[width=0.45\columnwidth]{../Figures/Pictures/doorway_BDPs}
	\\~\\
	\includegraphics[width=0.65\columnwidth]{../Figures/Pictures/roof_vent_BDPs.png}
	\caption[Bi-directional probe plus solid thermocouple arrays in East Structure.]{Bi-directional probe plus solid thermocouple array at the south exterior doorway (top) and roof vent (bottom) of the East Structure.}
	\label{fig:BDP_arrays}
\end{figure}

@@

\subsection{West Structure}
The first floor of the West Structure was instrumented with three bare-bead thermocouple arrays, two bi-directional probe plus solid thermocouple arrays, and one gas sample inlet pipe. The second floor was equipped with three bare-bead thermocouple arrays, four bi-directional probe plus solid thermocouple arrays, two total heat flux sensor pairs, and one gas sample inlet pipe. The location of the instrumentation in the West Structure is shown in Fig.~\ref{fig:west_instrumentation}.

\begin{figure}[!ht]
	\includegraphics[width=0.94\columnwidth]{../Figures/Floor_Plans/West_Structure_2nd_Floor_Dimensioned_Instrumentation}
	\\~\\
	\includegraphics[width=\columnwidth]{../Figures/Floor_Plans/West_Structure_1st_Floor_Dimensioned_Instrumentation}
	\caption[Locations and labels of instrumentation in the West Structure.]{Locations and labels of instrumentation in the second floor (top) and first floor (bottom) of the West Structure.}
	\label{fig:west_instrumentation}
\end{figure}

Every thermocouple array contained eight bare-bead thermocouples, each bi-directional probe and solid thermocouple array contained 8 probes, and gas concentrations were pulled through 9.5~mm (0.375~in) diameter stainless steel tubing located 1.2~m (4~ft) above the floor. Each pair of total heat flux sensors were located 1.0~m (3.3~ft) above the floor. The pair at A16 contained one sensor facing the ceiling and another facing the north side of the room, and the pair at A17 also contained one sensor facing the ceiling and another facing the stairway door.
\clearpage

@@

\subsection{Measurement Uncertainty}
\label{sec:Uncertainty}
There are different components of uncertainty in the reported length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity, and heat release rate measurements. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

@@

\subsubsection{Bi-Directional Probes}
Bi-directional probes and single thermocouples were used to measure the velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. A single bare-bead Chromel-Alumel (type K) thermocouple with a 1.0~mm (0.04~in) nominal diameter was co-located with each probe. The thermocouple wire was protected with a 3.2~mm (0.13~in) diameter inconel sheath. A gas velocity measurement study examining the doorway flow of pre-flashover compartment fires yielded expanded uncertainty measurements ranging from $\pm$0.14 to $\pm$0.22 for bi-directional probes similar to the ones described here~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments was estimated to be $\pm$18\%.

\subsubsection{Heat Release Rate}
A rotary type gas meter was used to measure the volume of propane provided to the gas burners. The manufacturer quantifies the maximum error as $\pm$2\%. The measurement was used in conjunction with the heat of combustion of propane to calculate the heat release rate of the fire. The total expanded uncertainty for the heat release rate obtained from this method was estimated to be $\pm13$\%. ***\textit{Still need to finish uncertainty estimate and include refs}***

@@

% ==========================
% = EXPERIMENTAL PROCEDURE =
% ==========================
\chapter{Experimental Procedure}
\label{chap:Experimental_Procedure}
A similar procedure was followed for all experiments described in this report. First, three propane burners were ignited in sequential order. Next, various actions, such as opening and closing doors and roof vents or turning on a fan, were performed to change the ventilation pattern within the structure. Finally, the burners were turned off and the fire was extinguished.

\section{East Structure Tests}
\label{sec:east_procedure}
Five different tests, Tests 2--6, were conducted in the East Structure. Table~\ref{table:HRR_East} lists the calculated heat release rate for each test.

Tests~2--4 followed an identical order of events. Fig.~\ref{fig:Tests_2-4_layout} includes a floor plan schematic and table of event times corresponding to the data files for each test. During Tests~2--4, the south exterior door was not able to completely close due to an obstruction caused by the hoses used to transport propane to the burners. So, when the south door was in the ``closed'' position, a gap of approximately 133~mm (5.25~in) was present between the door and its frame. During all other experiments, however, the south exterior doorway was not used to change the ventilation pattern of the structure and thus remained closed for the entirety of the test. For these tests, the hinged door was removed and gypsum board was used to fully close the doorway.

The procedures for Test~5 and Test~6 are outlined in a similar manner in Fig.~\ref{fig:east_test_5} and Fig.~\ref{fig:east_test_6}, respectively. Tests~5 and 6 involved performing a specific set of events three times sequentially. To avoid repetition of identical actions in the event time table, each set of events has been denoted as a ``sequence'' (abbreviated as ``seq.'' in the tables).

\begin{table}[!ht]
\caption{Calculated Heat Release Rate for East Structure Tests}
\begin{tabular}{lccccc}
 \toprule
 					& 	\multicolumn{5}{c}{\textbf{Heat Release Rate (kW)}} \\
\textbf{Test} 		&	\begin{tabular}{c} Corner \\ Burner On \end{tabular} 
					&	\begin{tabular}{c} Middle \\ Burner On \end{tabular}
					&	\begin{tabular}{c} Center \\ Burner On \end{tabular}
					& 	\begin{tabular}{c} Center \\ Burner Off \end{tabular}
					&   \begin{tabular}{c} Middle \\ Burner Off \end{tabular} \\
 \midrule
Test 2 				& 338 & 651	& 901 & 634 & 330 \\
Test 3 				& 326 & 650	& 905 & 631 & 326 \\
Test 4				& 351 & 654	& 952 & 652 & 337 \\
Test 5 - Seq. 1		& N/A & N/A	& 725 & N/A & N/A \\
Test 5 - Seq. 2		& N/A & N/A	& 728 & N/A & N/A \\
Test 5 - Seq. 3		& N/A & N/A	& 723 & N/A & N/A \\
Test 6 - Seq. 1		& N/A & N/A	& 722 & N/A & N/A \\
Test 6 - Seq. 2		& N/A & N/A	& 721 & N/A & N/A \\
Test 6 - Seq. 3		& N/A & N/A	& 714 & N/A & N/A \\
\bottomrule
\end{tabular}
\label{table:HRR_East}
\end{table}

@@

\section{West Structure Tests}
\label{sec:west_procedure}
Four different tests, Tests~22--25, were conducted in the West Structure. Tests~22 and 23 followed an identical procedure to change the ventilation. Fig.~\ref{fig:west_test_22} includes a floor plan schematic and table of event times corresponding to the data files for Tests~22 and 23. Similarly, Fig.~\ref{fig:west_test_24} includes a floor plan schematic and table of event times that outline the procedure followed during Tests~24 and 25.
